Making Sequential Programs Deterministic

### 2019-03-28 design.tex
Small changes

\begin{figure}
    $\infer[\rulename{any type}]{\wellformed{\kappa \  \beta}}{\kappa : \|Det / NonDet|}$

@@

    $\infer[\rulename{ordernondet collection}]{\wellformed{\kappa_c \  \beta_c \angles{\kappa_e \ \beta_e}}}{\kappa_c : \|OrderNonDet/Det| & \kappa_e \sqsubseteq \kappa_c & \beta_c \sqsubseteq \<Collection>}$

@@

%      $\infer[\rulename{ordernondet array}]{\wellformed{\kappa_e \ \beta_e \ \  \kappa_a []}}{\kappa_a : \|OrderNonDet/Det| & \kappa_e \sqsubseteq \kappa_a }$
%     
%     \bigskip
%     
%     $\infer[\rulename{nondet array}]{\wellformed{\|NonDet| \ \beta_e \ \  \|NonDet| []}}{}$
%     
%     \bigskip

    $\infer[\rulename{ordernondet map}]{\wellformed{\kappa_m \  \beta_m \angles{\kappa_k \ \beta_k, \kappa_v \ \beta_v}}}{\kappa_m: \|OrderNonDet/Det| & \kappa_k \sqsubseteq \kappa_m & \kappa_v \sqsubseteq \kappa_m & \beta_m \sqsubseteq \<Map>}$

@@

    \caption{Type well formedness}
\todo{The \rulename{any type} rule is not correct, because not every type is
  well-formed.  It needs an antecedent stating that $\beta$ is not a
  collection or map.}

\todo{The collection rules do not permit nested collections; similarly, the
  (currently commented out) array rules did not permit any two-dimensional
  array to be well-formed.  There is a bigger issue here.
  Consider the ``nondet collection'' rule.  $\beta_c$ is a class (we need
  to use a different symbol for it, probably $c$), but $\beta_e$ should be
  a full type, which might contain type qualifiers within it.  (Actually,
  $\|NonDet|\ \beta_e$ is a type, not $\beta_e$ itself.)  So, there should
  be an antecedent stating that $\|NonDet| \beta_e$ is well-formed, and
  for the ``ordernondet collection'' rule stating that $\kappa_e \beta_e$
  is well-formed.  Likewise, antecedents are needed for the map rules.
  Otherwise, these rules permit construction of $\<Map>\angles{\tau'}$
  where $\tau'$ is itself not well-defined.}

\todo{``$\kappa : \|Det / NonDet|$'' should be ``$\kappa = \|Det /
  NonDet|$'' or (better)
  ``$\kappa \in \{\|Det|, \|NonDet|\}$''.
  You could also express $\kappa_c : \|OrderNonDet/Det|$ as just
  $\kappa_c \sqsubseteq \|OrderNonDet|$ since $\|Det| \sqsubseteq \|OrderNonDet|$.
}

@@

\Cref{type-validity} shows the type validity\todo{The caption says
  ``type well formedness''.  Choose one term and use it consistently
  throughout.}
rules of \ourTypeSystem. 
As stated in the rule \rulename{any type}, the types \<Det> and \<NonDet>
can be written on any Java basetype.\todo{As noted in figure caption, that's
  not true, in some contexts such as a type argument.}

\begin{figure}
    \centering
    \begin{tabular}{|l|l|l|l|l|}
        \cline{3-5}
        \multicolumn{2}{c|}{~}  &  \multicolumn{3}{c|}{element type} \\ \cline{3-5}
        \multicolumn{2}{c|}{~}  & NonDet     & OrderNonDet & Det \\ \hline
        & NonDet      &   valid    &  invalid    & invalid  \\ \cline{2-5}
        Collection    & OrderNonDet &   invalid  &  valid      & valid  \\ \cline{2-5}
        & Det         &   invalid  &  invalid    & valid      \\ \hline
    \end{tabular}
    \caption{Valid Collection declarations.  The Collection's type qualifier
        must be a supertype or equal to the element type.}
    \label{fig:determinism-collections}
\end{figure}

@@

\OurTypeSystem supports three types of polymorphism:  type
polymorphism, basetype polymorphism, and qualifier polymorphism.
\todo{Give an example for each.}
\begin{itemize}
\item
In \theDeterminismCheckerImplementation,
type polymorphism is handled by Java's generics mechanism, which
\theDeterminismChecker fully supports, including class and method generics,
inference, etc.
Given the Java declaration
\begin{Verbatim}
<T> T identity(T arg) { return arg; }
\end{Verbatim}
the type of \<identity(arg)> is the same as the type of
\<arg>, and a programmer can write a type such as \codeid{List<@NonDet
  Integer>} (but see below for details about collections).
\item
Basetype polymorphism is enabled by writing a type qualifier on a use of a
type variable, which overrides the type qualifier at the instantiation
site.
For example, a \<choose> operation on sets could be defined in Java as
\begin{Verbatim}
class ChooseableSet<T> implements Set<T> {
  // return an arbitrary element from the set
  @NonDet T choose();
}
\end{Verbatim}
\item
Java does not provide a syntax that can be used for qualifier polymorphism,
so we use a special type qualifier name, \<@PolyDet>.
A qualifier-polymorphic method \<m> with signature $\forall \kappa . \kappa\ \<int> \times \<Det boolean> \rightarrow
\kappa\ \<String>$ would be written in our dialect of Java as
\begin{Verbatim}
@PolyDet String m(@PolyDet int, @Det boolean)
\end{Verbatim}
Each use of \<@PolyDet> stands for a use of the qualifier variable
$\kappa$, and there is no need to declare the qualifier variable $\kappa$.
\TheDeterminismChecker currently supports
qualifier
polymorphism on methods but not on classes.
Therefore, \<@PolyDet> may be written in methods (signatures and bodies)
but not on fields.
\end{itemize}

When a user writes a polymorphic qualifier on a method signature or a type parameter,
it indicates that it could resolve\todo{Should all occurrences of ``resolve'' be ``can be instantiated''} to any type in the type system depending on how it is used.
In \theDeterminismChecker, we define a polymorphic qualifier \<PolyDet>.
One of the most common locations for a polymorphic qualifier is at a method signature.
Consider the following declaration of addition:
\begin{verbatim}
@PolyDet int plus(@PolyDet int a, @PolyDet int b) { return a+b; }
\end{verbatim}
\todo{It has the standard meaning.  There are two ways to interpret this.}
There are two valid instantiations for this declaration:
\begin{enumerate}
    \item @NonDet int plus(@NonDet int a, @NonDet int b) { return a+b; }
    \item @Det int plus(@Det int a, @Det int b) { return a+b; }
\end{enumerate}
This indicates that method \<foo> can be called with arguments having any of the \codeid{Det or NonDet} type qualifiers.
%   (\<@OrderNonDet> is not allowed because it is invalid on primitive types).
 \<PolyDet> resolves to the least upper bound of
the actual types on arguments. For instance, if this method is called as \codeid{foo(@Det String arg1, @NonDet boolean arg2)}, \theDeterminismChecker resolves the method declaration as \codeid{@NonDet int foo(@NonDet String param1, @NonDet boolean param2)}
causing the return type to have the type qualifier \<NonDet>.

@@

\Cref{fig-poly-resolution} presents the rule for type checking of methods annotated with \<PolyDet> or its variants.
A polymorphic method type checks if every instantiation of \<PolyDet> type checks. If this method has \<PolyDet>$\uparrow$ or \<PolyDet>$\downarrow$
in its signature then the instantiation of \<PolyDet> at that location gets replaced by the appropriate type as shown in \cref{tab-poly-resolutions}.

@@

\begin{figure}
    $\infer[\rulename{poly resolution}]{\pi_r \ \beta_r\ \<m>(\overline{\pi_i \ \beta_i})\{\ \}}{\forall \kappa \in \{\|NonDet, OrderNonDet, Det|\} : \kappa \ \beta_r\ \<m>(\overline{K \ \beta_i})\{\ \} & \pi_r, \pi_i \in \{\|PolyDet|, \|PolyDet|\uparrow, \|PolyDet|\downarrow    \}}$
    \caption{Polymorphic resolution rule}
    \todo{I'm not sure what this is doing.  What code does it reject?}
    \todo{This suggests that \PolyDetUp is a qualifier, rather than
      $\uparrow$ being an operator on qualifiers.  We should clarify which
      it is.}
    \label{fig-poly-resolution}
\end{figure}

@@

\subsection{Discussion on defaulting rules}\label{defaulting}

\TheDeterminismChecker applies a default qualifier at each unannotated type
use (except uses of type variables, which already stand for a type that was
defaulted at the instantiation site).
This does not change the expressivity of the type system; it merely makes
the system more convenient to use.

In \theDeterminismChecker, we chose \<PolyDet> as the default for all 
method parameters and return types including constructors.
The rationale for this design choice was to not make \theDeterminismChecker too restrictive by making everything \<Det>.  This implies that
\theDeterminismChecker still allows some non-determinism as long as it doesn't break any determinism specifications explicitly written by the programmer.

\begin{verbatim}
@PolyDet int foo(@PolyDet int x) { ... }

void bar(@NonDet int random, @Det int value) {
    @NonDet int x = foo(random);      // No error.
    @Det int y = foo(value);          // No error.
    System.out.println(foo(random));  // Error.
}
\end{verbatim}
In the example above, the method \codeid{foo} is annotated with the \<PolyDet> type qualifier which allows
the caller method \codeid{bar} to invoke it with both \<Det> and \<NonDet> arguments. \TheDeterminismChecker flags an error 
only when the returned value is used in a way that breaks any deterministic specification (Printing a \<NonDet> value
breaks our determinism guarantees).

@@

\begin{figure}
    $\infer[\rulename{hashmap creation}]{\|new \ \kappa\ HashMap()|}{\kappa = \|OrderDet/NonDet|}$

@@

    $\infer[\rulename{treemap creation}]{\|new \ \kappa\ TreeMap()|}{\kappa = \|Det/NonDet|}$

@@

    $\infer[\rulename{hashset creation}]{\|new \ \kappa\ HashSet()|}{\kappa = \|OrderDet/NonDet|}$

@@

    $\infer[\rulename{treeset creation}]{\|new \ \kappa\ TreeSet()|}{\kappa = \|Det/NonDet|}$
    \caption{\<Map> and \<Set> creation rules}
    \label{fig-creation-rules}
\end{figure}

@@

\subsection{Determinism checker specific rules for precision}\label{precision}
We annotated the \codeid{equals()} method of the \<Set> class in the JDK as
\begin{verbatim}
PolyDet$\uparrow$ boolean equals(PolyDet Set<E> this, PolyDet Object o)
\end{verbatim}
While this is sound, it is imprecise in the case of \codeid{OrderNonDet Sets}. We added the following rule shown in \cref{fig-precision-rules} to our checker to improve precision.
\begin{figure}
    \todo{add line break}
    $\infer[\rulename{set precision}]{\Gamma \vdash \|z : Det|}{\Gamma \vdash \|x| : \|OrderNonDet| \ \|Set|\angles{ \kappa_{x} \ \beta_{x}}, \|y| : \|OrderNonDet| \ \|Set|\angles{ \kappa_{y} \ \beta_{y}} & \|z = x.equals(y)| & \angles{ \kappa_{x} \ \beta_{x}} \nsqsubseteq \angles{ \|OrderNonDet \ List| } & \angles{ \kappa_{y} \ \beta_{y}} \nsqsubseteq \angles{ \|OrderNonDet \ List| } }$

@@

Similarly, the \codeid{get()} method in the \<Map> class is annotated with polymorphic qualifiers and the rule $\sc{\|MAP \ PRECISION|}$ in \cref{fig-precision-rules} improves precision for \<OrderNonDet> receivers.

@@

\subsection{Extending \theDeterminismChecker}\label{extending-det-checker}
We have primarily focused on collections and annotated the corresponding classes in the JDK.
These qualifiers provide machine verifiable specifications of determinism properties.
Another source of nondeterminism that we haven't analyzed closely is the File system methods.
A program analysis developer interested in verifying determinism guarantees with respect to 
the File system can easily do so by annotating the parameter and return types of those methods in the JDK.
Alternatively, a user can write these qualifiers in a \textit{stub} file and supply it to the type checker. A \textit{stub}
only needs to contain method declarations along with their qualifiers.

>>>>

\begin{figure}
    $\infer[\rulename{noncollection}]
    {\wellformed{\kappa \  \beta}}
    {\kappa \in \{ \<Det>, \<NonDet> \} & \beta \not\sqsubseteq \<Collection> & \beta \not\sqsubseteq \<Map>}$

@@

    $\infer[\rulename{ordernondet collection}]
    {\wellformed{\kappa_c \  \beta_c \angles{\kappa_e \ \beta_e}}}
    {\kappa_c \sqsubseteq \|OrderNonDet| & \kappa_e \sqsubseteq \kappa_c &
      \beta_c \sqsubseteq \<Collection> & \wellformed{\kappa_e \ \beta_e}}$

@@

%      $\infer[\rulename{ordernondet array}]{\wellformed{\kappa_e \ \beta_e \ \  \kappa_a []}}{\kappa_a \sqsubseteq \|OrderNonDet| & \kappa_e \sqsubseteq \kappa_a }$
%     
%     \bigskip
%     
%     $\infer[\rulename{nondet array}]{\wellformed{\|NonDet| \ \beta_e \ \  \|NonDet| []}}{}$
%     
%     \bigskip

    $\infer[\rulename{ordernondet map}]
    {\wellformed{\kappa_m \  \beta_m \angles{\kappa_k \ \beta_k, \kappa_v \ \beta_v}}}
    {\kappa_m \sqsubseteq \|OrderNonDet|
      & \kappa_k \sqsubseteq \kappa_m
      & \kappa_v \sqsubseteq \kappa_m
      & \beta_m \sqsubseteq \<Map>
      & \wellformed{\kappa_k \ \beta_k}
      & \wellformed{\kappa_v \ \beta_v}
    }$

@@

    \caption{Type well formedness}
\todo{The \rulename{noncollection} rule is not correct, because not every type is
  well-formed.  It needs an antecedent stating that $\beta$ is not a
  collection or map.}

\todo{\todo{I think this will be OK, by redefining $\beta$.}
  The collection rules do not permit nested collections; similarly, the
  (currently commented out) array rules did not permit any two-dimensional
  array to be well-formed.  There is a bigger issue here.
  Consider the ``nondet collection'' rule.  $\beta_c$ is a class (we need
  to use a different symbol for it, probably $c$), but $\beta_e$ should be
  a full type, which might contain type qualifiers within it.  (Actually,
  $\|NonDet|\ \beta_e$ is a type, not $\beta_e$ itself.)  So, there should
  be an antecedent stating that $\|NonDet| \beta_e$ is well-formed, and
  for the ``ordernondet collection'' rule stating that $\kappa_e \beta_e$
  is well-formed.  Likewise, antecedents are needed for the map rules.
  Otherwise, these rules permit construction of $\<Map>\angles{\tau'}$
  where $\tau'$ is itself not well-defined.}

@@

\Cref{type-validity} shows the type validity\todo{The caption says
  ``type well formedness''.  Choose one term and use it consistently
  throughout.}
rules of \ourTypeSystem. 
As stated in the rule \rulename{noncollection}, the types \<Det> and \<NonDet>
can be written on any non-collection Java basetype.

\begin{figure}
    \centering
    \begin{tabular}{|l|l|l|l|l|}
        \cline{3-5}
        \multicolumn{2}{c|}{~}  &  \multicolumn{3}{c|}{Element type} \\ \cline{3-5}
        \multicolumn{2}{c|}{~}  & NonDet     & OrderNonDet & Det \\ \hline
                        & NonDet      &  valid     &  invalid    & invalid  \\ \cline{2-5}
        Collection type & OrderNonDet &  invalid   &  valid      & valid    \\ \cline{2-5}
                        & Det         &  invalid   &  invalid    & valid    \\ \hline
    \end{tabular}
    \caption{Valid Collection declarations.  The Collection's type qualifier
        must be a supertype or equal to the element type.}
    \label{fig:determinism-collections}
\end{figure}

@@

\OurTypeSystem supports three types of polymorphism:  type
polymorphism, basetype polymorphism, and qualifier polymorphism.
These apply to both classes and methods.
\begin{itemize}
\item
In \theDeterminismCheckerImplementation,
type polymorphism is handled by Java's generics mechanism, which
\theDeterminismChecker fully supports, including class and method generics,
inference, etc.
Given the Java declaration
\begin{Verbatim}
<T> T identity(T arg) { return arg; }
\end{Verbatim}
the type of \<identity(arg)> is the same as the type of
\<arg>, and a programmer can write a type such as \codeid{List<@NonDet
  Integer>} (but see below for details about collections).
\item
Basetype polymorphism is enabled by writing a type qualifier on a use of a
type variable, which overrides the type qualifier at the instantiation
site.
For example, a \<choose> operation on sets could be defined in Java as
\begin{Verbatim}
class ChooseableSet<T> implements Set<T> {
  // return an arbitrary element from the set
  @NonDet T choose();
}
\end{Verbatim}
\item
Java does not provide a syntax that can be used for qualifier polymorphism,
so we use a special type qualifier name, \<@PolyDet>.
A qualifier-polymorphic method \<m> with signature $\forall \kappa . \kappa\ \<int> \times \<Det boolean> \rightarrow
\kappa\ \<String>$ would be written in our dialect of Java as
\begin{Verbatim}
@PolyDet String m(@PolyDet int, @Det boolean)
\end{Verbatim}
Each use of \<@PolyDet> stands for a use of the qualifier variable
$\kappa$, and there is no need to declare the qualifier variable $\kappa$.
\TheDeterminismChecker currently supports
qualifier
polymorphism on methods but not on classes.
Therefore, \<@PolyDet> may be written in methods (signatures and bodies)
but not on fields.
\end{itemize}

When a user writes a polymorphic qualifier on a method signature or a type parameter,
it indicates that it could resolve\todo{Should all occurrences of ``resolve'' be ``can be instantiated''} to any type in the type system depending on how it is used.
In \theDeterminismChecker, we define a polymorphic qualifier \<PolyDet>.
One of the most common locations for a polymorphic qualifier is at a method signature.
Consider the following declaration of addition:
\begin{verbatim}
@PolyDet int plus(@PolyDet int a, @PolyDet int b) { return a+b; }
\end{verbatim}
\todo{It has the standard meaning.  There are two ways to interpret this.}
There are two valid instantiations for this declaration:
\begin{Verbatim}
  @NonDet int plus(@NonDet int a, @NonDet int b) { return a+b; }
  @Det int plus(@Det int a, @Det int b) { return a+b; }
\end{Verbatim}
This indicates that method \<foo> can be called with arguments having any of the \codeid{Det or NonDet} type qualifiers.
%   (\<@OrderNonDet> is not allowed because it is invalid on primitive types).
 \<PolyDet> resolves to the least upper bound of
the actual types on arguments. For instance, if this method is called as \codeid{foo(@Det String arg1, @NonDet boolean arg2)}, \theDeterminismChecker resolves the method declaration as \codeid{@NonDet int foo(@NonDet String param1, @NonDet boolean param2)}
causing the return type to have the type qualifier \<NonDet>.

@@

\Cref{fig-poly-resolution} presents the rule for type checking of methods annotated with \<PolyDet> or its variants.
A polymorphic method type checks if every instantiation of \<PolyDet> type checks. If this method has \PolyDetUp\ or \PolyDetDown\
in its signature then the instantiation of \<PolyDet> at that location gets replaced by the appropriate type as shown in \cref{tab-poly-resolutions}.

@@

\begin{figure}
    $\infer[\rulename{poly resolution}]{\pi_r \ \beta_r\ \<m>(\overline{\pi_i \ \beta_i})\{\ \}}{\forall \kappa \in \{\|NonDet, OrderNonDet, Det|\} : \kappa \ \beta_r\ \<m>(\overline{K \ \beta_i})\{\ \} & \pi_r, \pi_i \in \{\|PolyDet|, \PolyDetUp, \PolyDetDown    \}}$
    \caption{Polymorphic resolution rule}
    \todo{I'm not sure what this is doing.  What code does it reject?}
    \todo{This suggests that \PolyDetUp is a qualifier, rather than
      $\uparrow$ being an operator on qualifiers.  We should clarify which
      it is.}
    \label{fig-poly-resolution}
\end{figure}

@@

@@

\begin{figure}
    $\infer[\rulename{hashmap creation}]{\|new \ \kappa\ HashMap()|}{\kappa \in \{ \<OrderNonDet>, \<NonDet> \}}$

@@

    $\infer[\rulename{treemap creation}]{\|new \ \kappa\ TreeMap()|}{\kappa \in \{ \<Det>, \<NonDet> \}}$

@@

    $\infer[\rulename{hashset creation}]{\|new \ \kappa\ HashSet()|}{\kappa \in \{ \<OrderNonDet>, \<NonDet> \}}$

@@

    $\infer[\rulename{treeset creation}]{\|new \ \kappa\ TreeSet()|}{\kappa \in \{ \<Det>, \<NonDet> \}}$
    \caption{\<Map> and \<Set> creation rules}
    \label{fig-creation-rules}
\end{figure}

@@

\subsection{Determinism checker specific rules for precision}\label{precision}
We annotated the \codeid{equals()} method of the \<Set> class in the JDK as
\begin{verbatim}
PolyDet("down") boolean equals(PolyDet Set<E> this, PolyDet Object o)
\end{verbatim}
While this is sound, it is imprecise in the case of \codeid{OrderNonDet Sets}. We added the following rule shown in \cref{fig-precision-rules} to our checker to improve precision.
\begin{figure}
    \todo{add line break}
    $\infer[\rulename{set precision}]{\Gamma \vdash \|z : Det|}{\Gamma \vdash \|x| : \|OrderNonDet| \ \|Set|\angles{ \kappa_{x} \ \beta_{x}}, \|y| : \|OrderNonDet| \ \|Set|\angles{ \kappa_{y} \ \beta_{y}} & \|z = x.equals(y)| & \angles{ \kappa_{x} \ \beta_{x}} \nsqsubseteq \angles{ \|OrderNonDet \ List| } & \angles{ \kappa_{y} \ \beta_{y}} \nsqsubseteq \angles{ \|OrderNonDet \ List| } }$

@@

Similarly, the \codeid{get()} method in the \<Map> class is annotated with polymorphic qualifiers and the rule \rulename{map precision}\ in \cref{fig-precision-rules} improves precision for \<OrderNonDet> receivers.

@@

\subsection{Extending \ourTypeSystem}\label{extending-det-checker}
We have primarily focused on collections and annotated the corresponding classes in the JDK.
These qualifiers provide machine verifiable specifications of determinism properties.
Another source of nondeterminism that we haven't analyzed closely is the File system methods.
A program analysis developer interested in verifying determinism guarantees with respect to 
the File system can easily do so by annotating the parameter and return types of those methods in the JDK.
Alternatively, a user can write these qualifiers in a \textit{stub} file and supply it to the type checker. A \textit{stub}
only needs to contain method declarations along with their qualifiers.

### 2019-03-28 java-design.tex
Small changes

>>>>

\begin{itemize}
\item
In \theDeterminismCheckerImplementation,
type polymorphism is handled by Java's generics mechanism, which
\theDeterminismChecker fully supports, including class and method generics,
inference, etc.
Given the Java declaration
\begin{Verbatim}
<T> T identity(T arg) { return arg; }
\end{Verbatim}
the type of \<identity(arg)> is the same as the type of
\<arg>, and a programmer can write a type such as \codeid{List<@NonDet
  Integer>} (but see below for details about collections).
\item
Basetype polymorphism is enabled by writing a type qualifier on a use of a
type variable, which overrides the type qualifier at the instantiation
site.
For example, a \<choose> operation on sets could be defined in Java as
\begin{Verbatim}
class ChooseableSet<T> implements Set<T> {
  // return an arbitrary element from the set
  @NonDet T choose();
}
\end{Verbatim}
\item
Java does not provide a syntax that can be used for qualifier polymorphism,
so we use a special type qualifier name, \<@PolyDet>.
A qualifier-polymorphic method \<m> with signature $\forall \kappa . \kappa\ \<int> \times \<Det boolean> \rightarrow
\kappa\ \<String>$ would be written in our dialect of Java as
\begin{Verbatim}
@PolyDet String m(@PolyDet int, @Det boolean)
\end{Verbatim}
Each use of \<@PolyDet> stands for a use of the qualifier variable
$\kappa$, and there is no need to declare the qualifier variable $\kappa$.
\TheDeterminismChecker currently supports
qualifier
polymorphism on methods but not on classes.
Therefore, \<@PolyDet> may be written in methods (signatures and bodies)
but not on fields.
\end{itemize}

\subsection{Defaulting rules}\label{defaulting}

\TheDeterminismChecker applies a default qualifier at each unannotated type
use (except uses of type variables, which already stand for a type that was
defaulted at the instantiation site).
This does not change the expressivity of the type system; it merely makes
the system more convenient to use.

In \theDeterminismChecker, we chose \<PolyDet> as the default for all 
method parameters and return types including constructors.
The rationale for this design choice was to not make \theDeterminismChecker too restrictive by making everything \<Det>.  This implies that
\theDeterminismChecker still allows some non-determinism as long as it doesn't break any determinism specifications explicitly written by the programmer.

\begin{verbatim}
@PolyDet int foo(@PolyDet int x) { ... }

void bar(@NonDet int random, @Det int value) {
    @NonDet int x = foo(random);      // No error.
    @Det int y = foo(value);          // No error.
    System.out.println(foo(random));  // Error.
}
\end{verbatim}
In the example above, the method \codeid{foo} is annotated with the \<PolyDet> type qualifier which allows
the caller method \codeid{bar} to invoke it with both \<Det> and \<NonDet> arguments. \TheDeterminismChecker flags an error 
only when the returned value is used in a way that breaks any deterministic specification (Printing a \<NonDet> value
breaks our determinism guarantees).
